---
title: "Climate, Environment and Sustainability Research"
subtitle: '...Starting 2017'
author: ''
date: "`r Sys.Date()`"
fontsize: 20pt
mainfont: Helvetica
output:
  powerpoint_presentation: default
---


```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Goals

- Assess extent of climate-related research. 
- Create a baseline of climate-related research funding. 
- Disaggregate data based on important areas of focus.
- Develop a series of high impact case studies.
- Visualize findings and data.

## Questions

- Which are the primary funding bodies?
- What is the application success rate?
- How is research funding distributed among colleges/schools?
- What is the distribution of active projects?
- How is the research distributed around the world?
- What are the various research outcomes from projects?
- What are the major domains of research? 

```{r echo=FALSE, results='hide',message=FALSE}
gc(reset = T); rm(list = ls()) 
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, dplyr, readxl, purrr, ggVennDiagram, tidyverse, rnaturalearth, rnaturalearthdata)
```

```{r echo=FALSE, results='hide',message=FALSE}
# Paths, directories
datadir <- paste0(getwd(), "/data")
outdir <- paste0(getwd(), "/outputs")
```

```{r echo=FALSE, results='hide',message=FALSE, warning=FALSE}
# Reading projects from UKRI
projects_ukri <- read_csv(paste0(outdir, "/ukri_projects.csv")) %>% 
  mutate_at(vars(StartDate, EndDate), as.Date, format="%d/%m/%y") %>% 
  filter(StartDate >= "01/01/2017")

outcomes_ukri <- read_csv(paste0(outdir, "/ukri_outcomes.csv"))

projects_ero <- read_csv(paste0(outdir, "/ero_projects.csv")) %>% 
  mutate_at(vars(Application.Period.Start, Application.Period.End), as.Date, format="%y-%m-%d") %>% 
  filter(Academic.Year >= "2017")
```

## Source of Funding (2021)

```{r echo=FALSE, message=FALSE, fig.width = 10, fig.height = 5}
ero_projects_funding <- projects_ero %>% 
  filter(Academic.Year == "2021") %>% 
  group_by(Sponsor.Type) %>% 
  summarise(Number_projects = n()) %>% 
  mutate(Percentage = Number_projects/sum(Number_projects)*100) %>% 
  mutate_if(is.numeric, round, 1)

ggplot(ero_projects_funding, aes(x = reorder(Sponsor.Type, Percentage), y = Percentage, fill = Sponsor.Type)) +
  geom_bar(position="stack", stat="identity", width = 0.5) +
  #geom_text(aes(label=Percentage), hjust = -0.2)+
  labs(x = "",
       y = "Percent",
       title = "",
       caption = "Data source: ERO")+
  coord_flip()+
  theme_bw()+
  theme(plot.title = element_text(size = 12, family = "Helvetica", face = "bold"),
        text = element_text(size = 12, family = "Helvetica"),
        axis.title = element_text(face="bold", size = 14),
        axis.text.x=element_text(size = 12),
        axis.text.y=element_text(size = 12),
        strip.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "off")

```

## Application Success Rate

```{r echo=FALSE, message=FALSE, fig.width = 10, fig.height = 5}
ero_projects_funding <- projects_ero %>% 
  filter(!College %in% c("Corporate Services", "Information Services", "University Secretarys Group"),
         !Status %in% c("Pending")) %>% 
  group_by(Academic.Year, College, Status) %>% 
  summarise(Number_projects = n()) %>% 
  spread(Status, Number_projects) %>% 
  mutate(Success_rate = Successful/(Successful+Unsuccessful)*100) %>% 
  select(-c(Successful, Unsuccessful))


ggplot(ero_projects_funding, aes(factor(Academic.Year), Success_rate, group = College)) +
  geom_line(aes(color=College)) +
  geom_point(aes(color=College))+
  labs(x = "Year",
       y = "Percent",
       title = "",
       caption = "Data source: ERO")+
  theme_bw()+
  theme(plot.title = element_text(size = 14, family = "Helvetica", face = "bold"),
        text = element_text(size = 14, family = "Helvetica"),
        axis.title = element_text(face="bold", size = 16),
        axis.text.x=element_text(size = 14),
        axis.text.y=element_text(size = 14),
        strip.text = element_text(size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 14),
        legend.title=element_blank())+
  scale_colour_manual(values = c(`College of Arts, Humanities and Social Science` = "blue", 
                               `College of Medicine and Veterinary Medicine` = "skyblue",
                               `College of Science and Engineering` = "firebrick3"))

ggsave(paste0(outdir, "/ero_funding_sources.png"), width=1800, height=1000, units='px', dpi=150)
```


## Research Income (2021)

```{r echo=FALSE, message=FALSE, fig.width = 10, fig.height = 5}
ero_projects_income <- projects_ero %>% 
  filter(Academic.Year == "2021") %>% 
  filter(!College %in% c("Corporate Services", "Information Services", "University Secretarys Group")) %>% 
  group_by(College) %>% 
  summarise(Amount_funding = sum(Application.Total, na.rm = TRUE)/1000000) %>% 
  mutate_if(is.numeric, round, 1)

ggplot(ero_projects_income, aes(x = reorder(College, Amount_funding), y = Amount_funding, fill = College)) +
  geom_bar(position="stack", stat="identity", width = 0.5) +
  #geom_text(aes(label=Amount_funding), hjust = -0.2)+
  labs(x = "",
       y = "Amount in Millions",
       title = "",
       caption = "Data source: ERO")+
  coord_flip()+
  scale_y_continuous(labels = scales::label_dollar(prefix = "£"))+
  theme_bw()+
  theme(plot.title = element_text(size = 14, family = "Helvetica", face = "bold"),
        text = element_text(size = 14, family = "Helvetica"),
        axis.title = element_text(face="bold", size = 16),
        axis.text.x=element_text(size = 14),
        axis.text.y=element_text(size = 14),
        strip.text = element_text(size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "off")+
  scale_fill_manual(values = c(`College of Arts, Humanities and Social Science` = "blue", 
                               `College of Medicine and Veterinary Medicine` = "skyblue",
                               `College of Science and Engineering` = "firebrick3"))
```


## Source of Funding (Active projects)

```{r echo=FALSE, message=FALSE, fig.width = 10, fig.height = 5}
projects_funding <- projects_ukri %>% 
  group_by(FundingOrgName) %>% 
  summarise(Number_projects = n()) %>%
  mutate(Percentage = Number_projects/sum(Number_projects)*100) %>% 
  ungroup() %>% 
  mutate_if(is.numeric, round, 1)

ggplot(projects_funding, aes(x = reorder(FundingOrgName, Percentage), y = Percentage, fill = FundingOrgName)) +
  geom_bar(position="stack", stat="identity", width = 0.5) +
  #geom_text(aes(label=Percentage), hjust = -0.2)+
  labs(x = "",
       y = "Percent",
       title = "",
       caption = "Data source: UKRI")+
  coord_flip()+
  theme_bw()+
  theme(plot.title = element_text(size = 14, family = "Helvetica", face = "bold"),
        text = element_text(size = 14, family = "Helvetica"),
        axis.title = element_text(face="bold", size = 16),
        axis.text.x=element_text(size = 14),
        axis.text.y=element_text(size = 14),
        strip.text = element_text(size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "off")
```

## Distribution of active projects within colleges

```{r echo=FALSE, message=FALSE, fig.width = 10, fig.height = 5}
projects_distribution <- projects_ukri %>%
  filter(Status == "Active") %>% 
  group_by(College) %>% 
  summarise(Number_projects = n()) %>% 
  mutate(Percentage = Number_projects/sum(Number_projects)*100) %>% 
  ungroup() %>% 
  mutate_if(is.numeric, round, 1)

ggplot(projects_distribution, aes(x = reorder(College, Percentage), y = Percentage, fill = College)) +
  geom_bar(position="stack", stat="identity", width = 0.5) +
  #geom_text(aes(label=Percentage), hjust = -0.2)+
  labs(x = "",
       y = "Percent",
       title = "",
       caption = "Data source: UKRI")+
  coord_flip()+
  theme_bw()+
  theme(plot.title = element_text(size = 14, family = "Helvetica", face = "bold"),
        text = element_text(size = 14, family = "Helvetica"),
        axis.title = element_text(face="bold", size = 16),
        axis.text.x=element_text(size = 14),
        axis.text.y=element_text(size = 14),
        strip.text = element_text(size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "off")+
  scale_fill_manual(values = c(`College of Arts, Humanities & Social Sciences` = "blue", 
                               `College of Medicine & Veterinary Medicine` = "skyblue",
                               `College of Science and Engineering` = "firebrick3",
                               Other = "grey"))
```

## Categories of Projects

```{r echo=FALSE, message=FALSE, fig.width = 10, fig.height = 5}
projects_categories <- projects_ukri %>% 
  group_by(ProjectCategory) %>% 
  summarise(Number_projects = n()) %>% 
  mutate(Percentage = Number_projects/sum(Number_projects)*100) %>% 
  ungroup() %>% 
  mutate_if(is.numeric, round, 1)

ggplot(projects_categories, aes(x = reorder(ProjectCategory, Percentage), y = Percentage, fill = ProjectCategory)) +
  geom_bar(position="stack", stat="identity", width = 0.5) +
  #geom_text(aes(label=Percentage), hjust = -0.2)+
  labs(x = "",
       y = "Percent",
       title = "",
       caption = "Data source: UKRI")+
  coord_flip()+
  theme_bw()+
  theme(plot.title = element_text(size = 14, family = "Helvetica", face = "bold"),
        text = element_text(size = 14, family = "Helvetica"),
        axis.title = element_text(face="bold", size = 16),
        axis.text.x=element_text(size = 14),
        axis.text.y=element_text(size = 14),
        strip.text = element_text(size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "off")
```

## Geographic reach by College

```{r echo=FALSE, message=FALSE, fig.width = 10, fig.height = 5}
outcomes_reach_colleges <- outcomes_ukri %>% 
  group_by(College, `Geographical Reach`) %>% 
  summarise(Number_outcomes = n()) %>% 
  filter(!is.na(`Geographical Reach`)) %>% 
  mutate(Percentage = Number_outcomes/sum(Number_outcomes)*100) %>% 
  ungroup() %>% 
  mutate_if(is.numeric, round, 1)

ggplot(outcomes_reach_colleges, aes(x = reorder(College, Percentage), y = Percentage, fill = College)) +
  geom_bar(position="stack", stat="identity", width = 0.5) +
  #geom_text(aes(label=Percentage), hjust = -0.2)+
  labs(x = "",
       y = "Percent",
       title = "",
       caption = "Data source: UKRI")+
  coord_flip()+
  facet_wrap(~`Geographical Reach`)+
  theme_bw()+
  theme(plot.title = element_text(size = 14, family = "Helvetica", face = "bold"),
        text = element_text(size = 14, family = "Helvetica"),
        axis.title = element_text(face="bold", size = 16),
        axis.text.x=element_text(size = 14),
        axis.text.y=element_text(size = 14),
        strip.text = element_text(size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "off")+
    scale_fill_manual(values = c(`College of Arts, Humanities & Social Sciences` = "blue", 
                               `College of Medicine & Veterinary Medicine` = "skyblue",
                               `College of Science and Engineering` = "firebrick3",
                               Other = "grey"))
```

## Geographic reach by School

```{r echo=FALSE, message=FALSE, fig.width = 10, fig.height = 5}
outcomes_reach_schools <- outcomes_ukri %>% 
  group_by(College, School, `Geographical Reach`) %>% 
  summarise(Number_outcomes = n()) %>% 
  filter(!is.na(`Geographical Reach`)) %>% 
  group_by(College, School) %>% 
  mutate(Percentage = Number_outcomes/sum(Number_outcomes)*100) %>% 
  ungroup()

ggplot(outcomes_reach_schools, aes(x = reorder(School, Percentage), y = Percentage, fill = School)) +
  geom_bar(position="stack", stat="identity") +
  labs(x = "",
       y = "Percent",
       title = "",
       caption = "Data source: UKRI")+
  coord_flip()+
  facet_wrap(~`Geographical Reach`)+
  theme_bw()+
  theme(plot.title = element_text(size = 10, family = "Helvetica", face = "bold"),
        text = element_text(size = 10, family = "Helvetica"),
        axis.title = element_text(face="bold", size = 10),
        axis.text.x=element_text(size = 10),
        axis.text.y=element_text(size = 10),
        strip.text = element_text(size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "off")
```

```{r echo=FALSE, message=FALSE, eval = FALSE, fig.width = 10, fig.height = 5}
outcomes_reach_countries <- outcomes_ukri %>% 
  group_by(Country) %>% 
  summarise(Number_outcomes = n()) %>% 
  filter(!is.na(Country),
         Country != c("European Union (EU)", "Global"))

Country <- c("Austria", "Belgium", "Bulgaria", "Croatia", "Republic of Cyprus", "Czech Republic", "Denmark", "Estonia", "Finland", "France", "Germany", "Greece", "Hungary", "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg", "Malta", "Netherlands", "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", "Spain", "Sweden")
Number_outcomes <- rep(4, 27)
eu_countries <- data.frame(Country, Number_outcomes)

# Cleaning EU country list
outcomes_reach_countries <- rbind(outcomes_reach_countries, eu_countries) %>% 
  mutate(Country = recode(Country,
                          "Tanzania, United Republic of" = "Tanzania",
                          "Republic of Cyprus" = "Cyprus")) %>% 
  group_by(Country) %>% 
  summarise(Number_outcomes = n())

outcomes_reach_geo <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  rename(Country = admin) %>% 
  left_join(outcomes_reach_countries, by = "Country") %>% 
  mutate(Outcomes_global = 0) %>%
  replace(is.na(.), 0) %>% 
  rowwise() %>% 
  mutate(Foot_print = sum(Number_outcomes, Outcomes_global, na.rm = TRUE)) %>% 
  select(Country, Foot_print)

ggplot(data = outcomes_reach_geo) +
  geom_sf(aes(fill = Foot_print)) +
  scale_fill_gradient(low = "yellow", high = "red")+
  theme_bw()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        panel.border = element_blank(),
        legend.position = "off")
```

## Research outcomes

```{r echo=FALSE, message=FALSE, fig.width = 12, fig.height = 5}
outcomes_type_colleges <- outcomes_ukri %>% 
  group_by(College, `Outcome Type`) %>% 
  summarise(Number_outcomes = n()) %>% 
  #filter(!is.na(`Geographical Reach`)) %>% 
  mutate(Percentage = Number_outcomes/sum(Number_outcomes)*100) %>% 
  ungroup()

ggplot(outcomes_type_colleges, aes(x = reorder(`Outcome Type`, Percentage), y = Percentage, fill = `Outcome Type`)) +
  geom_bar(position="stack", stat="identity") +
  labs(x = "",
       y = "Percent",
       title = "",
       caption = "Data source: UKRI")+
  coord_flip()+
  facet_wrap(~College)+
  theme_bw()+
  theme(plot.title = element_text(size = 10, family = "Helvetica", face = "bold"),
        text = element_text(size = 10, family = "Helvetica"),
        axis.title = element_text(face="bold", size = 10),
        axis.text.x=element_text(size = 10),
        axis.text.y=element_text(size = 10),
        strip.text = element_text(size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "off")
```

## Domains of research

```{r echo=FALSE, message=FALSE, fig.width = 10, fig.height = 5}
ero_keyword_search <- read_csv(paste0(outdir, "/ero_keyword_search.csv"))

# Keyword categories
sheet_names <- excel_sheets(path=paste0(datadir, "/Keywords.xlsx"))[2:4] # Grab sheet names

# Loop through sheets
category_sheets <- lapply(sheet_names, function(x) {
  as.data.frame(read_excel(paste0(datadir, "/Keywords.xlsx"), sheet = x))})

# rename elements in list by sheet name
names(category_sheets) <- sheet_names

# Projects categorization
ero_projects_categorised <- list()

for (i in 1:length(category_sheets)){
  
  category_keys <- category_sheets[i]
  
  category_keys <- do.call(rbind.data.frame, category_keys)
  
  categorised_keys <- ero_keyword_search %>% 
    inner_join(category_keys, by = c("Keyword" = "Keywords")) %>%
    mutate(Category = Research_category) %>%
    select(Project_no) %>% 
    as.character()
  
  ero_projects_categorised[[i]] <- unlist(strsplit(categorised_keys, ","))
  
}

names(ero_projects_categorised) <- sheet_names

ggVennDiagram(ero_projects_categorised, color = 1, lwd = 0.7, lty = 1, label_alpha = 0, label_percent_digit = 1, label = "percent", label_size = 5) + 
  scale_fill_distiller(palette = "Reds", direction = 1) +
  theme_bw()+
  theme(plot.title = element_text(size = 16, family = "Helvetica", face = "bold"),
        text = element_text(size = 16, family = "Helvetica"),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        strip.text = element_text(size = 16),
        panel.border=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background=element_blank(),
        legend.position = "off")
```
